Using SMT engine to generate Symbolic Automata\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-01-18 2017-Xudong-SMT/FM-2018-extended.tex
proofreading of BIP part

\documentclass{lncs/llncs}

@@

\usepackage{soul}               % Highlighting: \hl, \so, \ul
\soulregister{\st}{8}
\soulregister\cite7
\soulregister\citep7
\soulregister\secn7
\soulregister\ref7
\soulregister\pageref7
\soulregister{\em}{0}
\soulregister{\bf}{0}
\soulregister{\fig}{7}
\soulregister{\ie}{0}
\soulregister{\etc}{0}
\soulregister{\eg}{0}
\soulregister{\wrt}{0}
\soulregister{\resp}{0}
\soulregister{\cf}{0}
\soulregister{\mdash}{0}
\soulregister{\ndash}{0}

@@

\newcommand{\noteSB}[2][color=green!40, size=\tiny]{\todo[#1]{{\bf Note: } {#2}}}
\newcommand{\noteInSB}[2][inline,color=green!40]{\todo[#1]{{\bf Note: } {#2}}}

@@

\paragraph{Structure.}
In section
\ref{section:pnets} we give a description and a formal definition of
the pNet model, as found in previous publications. Then in
\ref{section:examples} we present a simple example that will
illustrate the rest of the paper.
Section \ref{section:op-semantics} recalls the operational semantics
of pNet, including its structural rules.
Section \ref{section:implementation} explains in details our
implementation within the Vercors platform, and shows the full result of
the semantic computation on the running example.
Finally we conclude and discuss perspectives in section
\ref{section:conclusion}.

@@

@@

In this context, the notion of \emph{architecture} was proposed
in~\cite{AttieBBJS16-architectures-faoc} as a mechanism for ensuring
correctness by construction during the design of BIP models.
Architecutres can be viewed as operators transforming BIP models.
They formalise design patterns, which enforce global properties
characterising the coordination among the components of the system.
The architecture-based design process in BIP takes as input a set of
components providing basic functionality of the system and a set of
temporal properties that must be enforced in the final system.  For
each property, a corresponding architecture is identified\,\footnote{%
%
  The current approach relies on taxonomies of predefined
  architectures.  Future work may include automatic synthesis of
  architectures from, \eg CTL or a similar logic. 
%
} and applied to the model, thereby potentially introducing additional
coordinator components and modifying the connectors that define
synchronisation patterns among ports of components.
In~\cite{AttieBBJS16-architectures-faoc}, it was shown that
application of architectures is compositional \wrt safety properties,
\ie when several architectures are applied, each enforcing a safety
property, the resulting system satisfies their conjunction.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.8\columnwidth]{XFIG/FailureTimerPNET}
  \caption{pNet encoding of the FailureTimer architecture}
  \label{schema:ArchFailure:pNet}
\end{figure}

\noteSB{SV2 should not be closed;\\SV0, SV1: $\cup \rightarrow \lor$}\hl{%
%
Figure~\ref{schema:ArchFailure:pNet} shows a pNet encoding of the
above Failure Monitor architecture.  This encoding is structural: each
coordinator component is encoded as a pLTS, the operand
component\mdash as a hole; connectors of the BIP model are encoded as
synchronisation vectors.  Each connector that does not involve
triggers is trivially encoded by a synchronisation vector comprising
the same ports.  In order to encode the semantics of the connectors
involving triggers, we 1)~in the pLTS encoding the coordinator
components, add loop transitions to ensure that all ports involved in
such connectors are enabled in all states, 2)~associate a Boolean
value to each of these ports: the original transitions carry the value
$\true$ (\eg $\cstate{0} \goesto{\PortFail(\true)} \cstate{1}$), the
added loops carry the value $\false$ (\eg $\cstate{2}
\goesto{\PortFail(\false)} \cstate{2}$), 3)~add to the corresponding
synchronisation vectors the Boolean predicate encoding the connector
structure.  For example, $SV0$ encodes the connector discussed above:
the predicate $b1 = b2 \land b1 \lor b2 \Rightarrow b0$ means that the
``true'' transitions $\NameCtrl.\PortFail$ and $\NameTimer.\PortStart$
can only fire together ($b1 = b2$) and whenever one of them fires,
$\NameOprnd.\PortFail$ must fire also ($b1 \lor b2 \Rightarrow b0$).
This encoding can be systematically obtained for any hierarchical BIP
connector~\cite{BliSif10-causal-fmsd}.
%
}
\noteSB{
%
  Eric, could you please check my pNet terminology in this paragraph?
%
}
%
Although, for the sake of brevity, this encoding does not translate
the maximal progress priority, it can be easily extended to do so, by
introducing an additional Boolean variable for each relevant port.

@@

\begin{example}\emph{An open transition.}
  This transition is generated by application of vector \texttt{SV0},
  synchronizing the initial actions of pLTSs \texttt{Failure} and
  \texttt{Timer}, with an action of the hole \texttt{B} equal to
  \texttt{fail}. The local variable \texttt{t} of the \texttt{Timer}
  is assigned to its initial value \texttt{Max}. \TODO{may be the
    place to mention the ``fresh variable'' names, rather than in th
    eZ3 syntax figure}
  \label{OT:lotos-composed}
        \begin{mathpar}
	\inferrule*[myfraction=\reddottedrule]
	{\{s0~{\xrightarrow{fail(true)}} ~s1, t0~{\xrightarrow{start(true)}} ~t1\}
	  \{\xrightarrow{hb}\}, hb=fail(true) \land v=\underline{fail}, \{t := Max\}
        }
	{\ostate{s0,t0} \xrightarrow {v} \ostate{s1,t1}}
	\end{mathpar}
\end{example}
%
%The open transition $ot_2$ involves the possible actions from the process $P$ and prefix expression $acc(x);Q$. The hole behavior, represented by variable $hb_{12}$ must be equal to the exit action $\delta(x)$, and the accept event $acc(x)$ will capture the same value $x$.
%
%The transition in the conclusion of this rule starts in
%state $\ostate{00}$, where both $C_1$ and $C_2$ are in their initial
%states. It ends in a different structured state, because in $ot_2$
%both pLTS conduct transitions, showing that the control has gone from
%processes $P$ to $(acc(x);Q$, consuming actions $\delta(x)$ and
%$acc(x)$. Controller $C_1$ has only one state, but its state variable
%$s_0$ has changed to 1.

@@

\section{Variable managment}
The variables in each synchronisation vector are considered local: for
a given pNet expression, we must have fresh local variables for each
occurrence of a vector (= each time we instantiate rule
Tr2). Similarly the state variables of each copy of a given pLTS in
the system, must be distinct, and those created for each application
of Tr2 have to be fresh and all distinct. This will be implemented
within the open-automaton generation algorithm, e.g. using name
generation using a global counter as a suffix.

@@

\section{Full examples}

\subsection{Full input for the FailureTimer architecture}
\subsection{Full output for the FailureTimer open automaton}

>>>>

\documentclass{lncs/llncs}
\usepackage{etex}

@@

\usepackage{soul}               % Highlighting: \hl, \so, \ul

\soulregister{\st}{8}
\soulregister\cite7
\soulregister\citep7
\soulregister\secn7
\soulregister\ref7
\soulregister\pageref7
\soulregister{\em}{0}
\soulregister{\bf}{0}
\soulregister{\fig}{7}
\soulregister{\ie}{0}
\soulregister{\etc}{0}
\soulregister{\eg}{0}
\soulregister{\wrt}{0}
\soulregister{\resp}{0}
\soulregister{\cf}{0}
\soulregister{\mdash}{0}
\soulregister{\ndash}{0}

@@

\newcommand{\noteSB}[2][color=green!40, size=\tiny]{\todo[#1]{{\bf
      Note: } {#2}}}
\newcommand{\noteEM}[2][color=blue!40, size=\tiny]{\todo[#1]{{\bf Eric: } {#2}}}
\newcommand{\noteInSB}[2][inline,color=green!40]{\todo[#1]{{\bf Note: } {#2}}}

@@

\paragraph{Structure.}
In section
\ref{section:pnets} we give a description and a formal definition of
the pNet model, as found in previous publications. Then in section
\ref{section:examples} we present our use-case, based on an
industrial-inspired study of the control software of nano-satellites,
specified using the BIP language. 
Section \ref{section:op-semantics} recalls briefly the operational semantics
of pNet.
Section \ref{section:implementation} explains in details the algorithm
used to compute this semantics, including the interaction with Z3, and
shows the full result of the semantic computation on the running example.
Finally we conclude and discuss perspectives.
\ref{section:conclusion}.

@@

In Fig.~\ref{schema:ArchFailure:pNet} we show examples of these
constructs, with 2 pLTSs and one pNet node encoding our use-case.

@@

In this context, the notion of \emph{architecture} was proposed
in~\cite{AttieBBJS16-architectures-faoc} as a mechanism for ensuring
correctness by construction during the design of BIP models.
Architectures can be viewed as operators transforming BIP models.
They formalise design patterns, which enforce global properties
characterising the coordination among the components of the system.
The architecture-based design process in BIP takes as input a set of
components providing basic functionality of the system and a set of
temporal properties that must be enforced in the final system.  For
each property, a corresponding architecture is identified\,\footnote{%
%
  The current approach relies on taxonomies of predefined
  architectures.  Future work may include automatic synthesis of
  architectures from, \eg CTL or a similar logic. 
%
} and applied to the model, thereby potentially introducing additional
coordinator components and modifying the connectors that define
synchronisation patterns among ports of components.
In~\cite{AttieBBJS16-architectures-faoc}, it was shown that
application of architectures is compositional \wrt safety properties,
\ie when several architectures are applied, each enforcing a safety
property, the resulting system satisfies their conjunction.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.9\columnwidth]{XFIG/FailureTimerPNET}
  \caption{pNet encoding of the FailureTimer architecture
  \noteEM{A little too small for me, my eyes are getting tired...}}
  \label{schema:ArchFailure:pNet}
\end{figure}

\noteSB{SV2 should not be closed;\\SV0, SV1: $\cup \rightarrow \lor$}
%\hl{%
%
Figure~\ref{schema:ArchFailure:pNet} shows a pNet encoding of the
above Failure Monitor architecture.  This encoding is structural: each
coordinator component is encoded as a pLTS, the operand
component\mdash as a hole; connectors of the BIP model are encoded as
synchronisation vectors.  Each connector that does not involve
triggers is trivially encoded by a synchronisation vector comprising
the same ports.  In order to encode the semantics of the connectors
involving triggers, we 1)~in the pLTS encoding the coordinator
components, add loop transitions to ensure that all ports involved in
such connectors are enabled in all states, 2)~associate a Boolean
value to each of these ports: the original transitions carry the value
$\true$ (\eg $\cstate{0} \goesto{\PortFail(\true)} \cstate{1}$), the
added loops carry the value $\false$ (\eg $\cstate{2}
\goesto{\PortFail(\false)} \cstate{2}$), 3)~add to the corresponding
synchronisation vectors the Boolean predicate encoding the connector
structure.  For example, $SV0$ encodes the connector discussed above:
the predicate $(b1=b2) \land (b1\lor b2 \Rightarrow b0)$ means that the
``true'' transitions $\NameCtrl.\PortFail$ and $\NameTimer.\PortStart$
can only fire together ($b1 = b2$) and whenever one of them fires,
$\NameOprnd.\PortFail$ must fire also ($b1 \lor b2 \Rightarrow b0$).
This encoding can be systematically obtained for any hierarchical BIP
connector~\cite{BliSif10-causal-fmsd}.
%
%}
\noteSB{
%
  Eric, could you please check my pNet terminology in this paragraph?
%
}
\noteEM{Done, it's ok.}
%
Although, for the sake of brevity, this encoding does not translate
the maximal progress priority, it can be easily extended to do so, by
introducing an additional Boolean variable for each relevant port.

@@

\begin{example}\emph{An open transition.}
  This transition is generated by application of vector \texttt{SV0},
  synchronizing the initial actions of pLTSs \texttt{Failure} and
  \texttt{Timer}, with an action of the hole \texttt{B} equal to
  \texttt{fail}. The local variable \texttt{t} of the \texttt{Timer}
  is assigned to its initial value \texttt{Max}.
In fact this OT is displayed in a simplied, human readable form. The
raw output of the tool is slightly more difficult to read, in
particular it uses structured names for the fresh variables generated
by the algorithm, allowing tracability of the result (see
Appendix~\ref{appendix:varaibles}). A full output of 
the use-case is provided in Appendice~\ref{appendix:fullOutput}.

  \label{OT:lotos-composed}
        \begin{mathpar}
	\inferrule*[myfraction=\reddottedrule]
	{\{s0~{\xrightarrow{fail(true)}} ~s1, t0~{\xrightarrow{start(true)}} ~t1\}
	  \{\xrightarrow{hb}\}, hb=fail(true) \land v=\underline{fail}, \{t := Max\}
        }
	{\ostate{s0,t0} \xrightarrow {v} \ostate{s1,t1}}
	\end{mathpar}
\end{example}
%
%The open transition $ot_2$ involves the possible actions from the process $P$ and prefix expression $acc(x);Q$. The hole behavior, represented by variable $hb_{12}$ must be equal to the exit action $\delta(x)$, and the accept event $acc(x)$ will capture the same value $x$.
%
%The transition in the conclusion of this rule starts in
%state $\ostate{00}$, where both $C_1$ and $C_2$ are in their initial
%states. It ends in a different structured state, because in $ot_2$
%both pLTS conduct transitions, showing that the control has gone from
%processes $P$ to $(acc(x);Q$, consuming actions $\delta(x)$ and
%$acc(x)$. Controller $C_1$ has only one state, but its state variable
%$s_0$ has changed to 1.

@@

\section{Variable managment}
\label{appendix:varaibles}
The variables in each synchronisation vector are considered local: for
a given pNet expression, we must have fresh local variables for each
occurrence of a vector (= each time we instantiate rule
Tr2). Similarly the state variables of each copy of a given pLTS in
the system, must be distinct, and those created for each application
of Tr2 have to be fresh and all distinct. This will be implemented
within the open-automaton generation algorithm, e.g. using name
generation using a global counter as a suffix.

@@

\section{Full output for the FailureTimer open automaton}
\label{appendix:fullOutput}
